---
title: "PA1_template.rmd"
author: "Ammar"
date: "9/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#### Load the data

```{r }
data <- read.csv("activity.csv")
```

#### Mean of the total number of steps taken per day
```{r data}
step_by_day<- aggregate(steps~date, data,sum)
hist(step_by_day$steps,main="Total steps each day",col="green", xlab ="Number of steps")
rmean<- mean(step_by_day$steps)
rmean
rmedian<- median(step_by_day$steps)
rmedian
```
The mean of the number of steps per day is `r rmean`.
The median of the number of steps per day is `r rmedian`.

#### Average daily activity pattern

```{r }
steps_by_interval<- aggregate(steps~interval, data,mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps", main="Average Number of per Day by Interval")
max_interval<- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```
The interval with most steps is `r max_interval`.

#### Imputing missing values
##### Calculate total number of missing values in the dataset

```{r}
NATotal<- sum(!complete.cases(data))
NATotal
```
Total number of missing values are `r NATotal`.

##### Using Mean for missing values
```{r}
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
    obs <- data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```

##### Creating a new data set including the imputed values
```{r}
new_activity<- data
new_activity$steps<- fillNA
```

#### Histogram for total number of steps taken each day after imputing

```{r}
StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
#Create Histogram to show difference. 
hist(step_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps", add=TRUE)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "green"), lwd=10)
```


#### calculate the mean and median
```{r}
rmeantotal<- mean(StepsTotalUnion$steps)
rmeantotal
rmediantotal<- median(StepsTotalUnion$steps)
rmediantotal
```

#### Do this differ from the non imputed value?
```{r}
rmeandiff<- rmeantotal- rmean
rmeandiff
rmediandiff<- rmediantotal- rmedian
rmediandiff
```
##### The mean is same for both, while median has a small variance.


#### If there is difference in activity patterns between weekdays and weekends?
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```

Created a plot to compare and contrast number of steps between the week and weekend. There is a higher peak earlier on weekdays, and more overall activity on weekends.













